---
title: 'Lab 01 - Functions and Vectors'
author: "Muge Kosar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
---

```{r include = FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

<br>

# Data

Data dictionary: [Syracuse city tax parcel data](http://data.syrgov.net/datasets/1b346804e1364a5eb85ccb53302e3c91_0) 

```{r}

URL <- "https://raw.githubusercontent.com/DS4PS/Data-Science-Class/master/DATA/syr_parcels.csv"
dat <- read.csv( URL, stringsAsFactors=FALSE )

```

```{r, echo = FALSE}

head( dat, 10 )         # Preview a dataset

```

```{r, echo = FALSE}

head( dat$owner, 10 )   # Preview a variable ("vector")

```

<br>

# Solutions

## Question 1

*How many tax parcels are in Syracuse, NY?* 

**ANSWER:** There are **41502** tax parcels in Syracuse, NY. 


```{r}

nrow(dat)       # Find dataset dimensions which represent tax parcels in Syracuse, NY

```

<br>

## Question 2

*How many acres of land are in Syracuse, NY?*

**ANSWER:** There are **12510.49** acres of land in Syracuse, NY.

```{r}

# Summarize numeric acres vector and ignore missing values which will count acres of land in Syracuse, NY

sum(dat$acres,na.rm = TRUE)   

```

<br>

## Question 3

*How many vacant buildings are there in Syracuse, NY?*

**ANSWER:** There are **1888** vacant buildings in Syracuse, NY.

```{r}

# Summarize vacantbuil vector and ignore missing values which will count vacant buildings in Syracuse, NY

sum(dat$vacantbuil,na.rm = TRUE)  

```


## Question 4

*What proportion of parcels are tax exempt?*

**ANSWER:** **10.70069%** of parcels are tax exempt.

```{r}

# Find number of TRUEs on tax.exempt variable & get proportion of TRUEs which gives proportion of tax-exempt parcels

sum(dat$tax.exempt,na.rm = TRUE) / length(dat$tax.exempt)      

# Alternative option which gives same result but with mean function

mean(dat$tax.exempt,na.rm = TRUE)

```

<br>

## Question 5

*Which neighborhood contains the most tax parcels?*

**ANSWER:** **Eastwood** contains the most tax parcels.

```{r}

table(dat$neighborhood)       # Find  the frequency of neighborhood variable

# (Optional) Find the maximum frequency of tax parcels and check name of neighborhood to narrow result

max(table(dat$neighborhood),na.rm = TRUE)           # 

```

<br>

## Question 6

*Which neighborhood contains the most vacant lots?*

**ANSWER:** **Near Westside** containts the most vacant lots.

```{r}

# Cross-tabulation of neighborhood and land_use variables

table(dat$neighborhood,dat$land_use)          

```
