---
title: "Lab 06: Dataset Joins"
author: "Muge Kosar"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r include = FALSE, cache = TRUE}

library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(pander)
library(scales)

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10)     # Global knitr options

options(scipen = 999)                     # Disable sci. notation

`%!in%` <- Negate(`%in%`)                 # Custom filter function

url <- "https://www.onetcenter.org/dl_files/database/db_26_3_text/"

dbs <- c("Skills",
         "Knowledge",
         "Abilities",
         "Interests",
         "Job%20Zones",
         "Work%20Values",
         "Work%20Styles",
         "Occupation%20Data",
         "Alternate%20Titles",
         "Technology%20Skills",
         "Job%20Zone%20Reference",
         "Education%2C%20Training%2C%20and%20Experience",
         "Education%2C%20Training%2C%20and%20Experience%20Categories")

txt <- paste0(url, dbs, ".txt")

all <- lapply(txt, read_delim)

names(all) <- gsub(pattern = "Education%2C%20Training%2C%20and|%20",
                   replacement = "",
                   x = dbs)

for(i in 1:length(all)) assign(names(all)[i], all[[i]])   # Create objects

rm(all)

```

<br>

## Assignment Overview

Follow the instructions below.

<br>

The following objects have been created for you from the [O*NET 26.3 Database](https://www.onetcenter.org/database.html). 

1. `Abilities` measures innate or natural abilities of occupational incumbents
2. `AlternateTitles` lists over 53,000 alternate names for O*NET occupations
3. `Experience` measures various kinds of experience, albeit encoded
4. `ExperienceCategories` decodes measured experience
5. `Interests` measures occupations with the Holland Code (RIASEC) Test  
6. `JobZoneReference` decodes classes of career development
7. `JobZones` broad classes of career development for each occupation; encoded
8. `Knowledge` defines areas of expertise and importance
9. `OccupationData` lists the occupation SOC, title, and description
10. `Skills` measures learned abilities of occupational incumbents
11. `TechnologySkills` lists all relevant technology skills
12. `WorkStyles` measures various performance indicators
13. `WorkValues` measures important occupation qualities 

<br>
<br>
<br>

## Part 1: Basic Join Challenges

The following challenges require no more then 2-3 tables and use the same column 
names across all tables. Use basic `dplyr` verbs to answer specific questions.

<br>

**(1) Question:** Which 10 occupation titles (`OccupationData$Title`) have the 
highest number of "hot technologies" (`TechnologySkills$HotTechnology`)?

```{r cache=TRUE}

OccupationData %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Hot Technology`) %>%
  filter(`Hot Technology` == "Y") %>%
  group_by(Title) %>%
  summarize(`Hot Technology`=n()) %>%
  arrange(-`Hot Technology`) %>%
  head(10)

```

<br>

**(2) Question:** Which 10 occupation titles (`OccupationData$Title`), with more 
than 15 unique technologies, have the highest proportion (%) of "hot technologies" 
(`TechnologySkills$HotTechnology`)?

```{r cache = TRUE}

OccupationData %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Hot Technology`) %>%
  filter(`Hot Technology` == "Y")%>%
  group_by(Title) %>%
  summarize(Total_tech = n()) %>%
  ungroup() %>%
  mutate(Prop = Total_tech / sum(Total_tech),
         Prop = percent(Prop))  %>%
  arrange(-Total_tech) %>%
  head(10)

```

<br>

**(3) Question:** Which unique occupation titles (`OccupationData$Title`) have 
"R" listed as a technology example (`TechnologySkills$Example`)? 

```{r cache = TRUE}

OccupationData %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  select(Title,Example) %>%
  filter(Example == "R") 


```

<br>

**(4) Question:** Which alternate occupation titles (`AlternateTitles`) have "R" 
listed as a technology example (`TechnologySkills$Example`)? 

```{r cache = TRUE}

AlternateTitles %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  select(`Alternate Title`,Example) %>%
  filter(Example == "R") 

```

<br>

**(5) Question:** Which top 25 occupation titles (`OccupationData$Title`) have 
the highest level of "Independence" (`Element Name`) according to `WorkValues`?

```{r eval=FALSE}

OccupationData %>% 
  left_join(WorkValues, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Element Name`, `Data Value`) %>%
  filter(`Element Name` == "Independence") %>%
  arrange(-`Data Value`) %>%
  head(25)

```

<br>

**(6) Question:** Which top 25 occupation titles (`OccupationData$Title`) have 
the highest level of "Persistence" (`Element Name`) according to `WorkStyles`?

```{r eval=FALSE}

OccupationData %>% 
  left_join(WorkStyles, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Element Name`, `Data Value`) %>%
  filter(`Element Name` == "Persistence") %>%
  arrange(-`Data Value`) %>%
  head(25)
```

<br>

**(7) Question:** According to Holland Codes, "Realistic" describes people who 
like to work with things, requiring "motor coordination, skill, and strength".
Which top 20 occupation titles (`OccupationData$Title`) have the highest rating 
in "Realistic" scores (`Element Name`) according to `Interests`?

```{r cache = T}

OccupationData %>% 
  left_join(Interests, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Element Name`, `Data Value`) %>%
  filter(`Element Name` == "Realistic") %>%
  arrange(-`Data Value`) %>%
  head(20)

```

<br>

**(8) Question:** Again, per Holland Codes, "Investigative" describes people who 
like to work with ideas, preferring observation over action and facts over 
feelings. Which top 20 occupation titles (`OccupationData$Title`) have the 
highest rating in "Investigative" scores (`Element Name`) according to `Interests`?

```{r cache = T}

OccupationData %>% 
  left_join(Interests, 
            by = "O*NET-SOC Code") %>%
  select(Title,`Element Name`, `Data Value`) %>%
  filter(`Element Name` == "Investigative") %>%
  arrange(-`Data Value`) %>%
  head(20)

```

<br>

**(9) Question:** Job Zones loosely classify how advanced one typically must be 
as an occupational incumbent. How many unique occupation titles 
(`OccupationData$Title`) are classified in each `Job Zone` (`JobZones`)? 

```{r eval=FALSE}

OccupationData %>% 
  left_join(JobZones, 
            by = "O*NET-SOC Code") %>%
  select(Title, `Job Zone`) %>%
  group_by(`Job Zone`) %>%
  summarize(Title=n())
```

<br>

**(10) Question:** Of occupation titles (`OccupationData$Title`) that list "R" 
as the `Example` in `TechnologySkills`, what proportion is in each Job Zone 
(`JobZones`)?

```{r eval=FALSE}

OccupationData %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  filter(Example == "R") %>%
  left_join(JobZones,
             by = "O*NET-SOC Code") %>%
  select(Title,`Job Zone`) %>%
  group_by(`Job Zone`) %>%
  summarize(Title=n()) %>%
  ungroup() %>%
  mutate(prop = Title / sum(Title))

```

<br>
<br>
<br>

## Part 2: Joins & Salary Predictions

The following challenges require 3-4 merged tables and incorporate data from the 
U.S. Bureau of Labor Statistics' 2021
[Occupational Employment and Wage Statistics](https://www.bls.gov/oes/current/oes_nat.htm)
which, like other federal agencies, uses unique SOC codes for each occupation.

<br>

The following commands will read in BLS data and rename and reformat the columns 
to match O*NET.

```{r cache = TRUE, message = FALSE, warning = FALSE}

url <- paste0("https://raw.githubusercontent.com/DS4PS/ays-r-cod",
              "ing-sum-2022/main/labs/bls_occupation_salaries.csv")

Salaries <- read_csv(url, 
                     col_select = c("O*NET-SOC Code" = "OCC_CODE", 
                                    "Title" = "OCC_TITLE", 
                                    "Median" = "A_MEDIAN")) %>% 
  mutate(`O*NET-SOC Code` = gsub(x = `O*NET-SOC Code`, 
                                 pattern = "$", 
                                 replacement = "\\.00"),
         Median = gsub(x = Median, 
                       pattern = ",", 
                       replacement = ""),
         Median = as.numeric(Median))

```

<br>

**Note:** There is one key distinction between O*NET and BLS SOC codes. The 
[2019 O*NET Taxonomy](https://www.onetcenter.org/taxonomy.html) features even 
more detailed profiles than typical federal agencies. Some more detailed O*NET 
occupation profiles will not successfully merge with BLS occupations.

<br>

**(1) Question:** Join the O*NET `OccupationData` and BLS `Salaries` tables. 
Then, filter out any occupations with an `NA` value in variable `Median`.
Store this new table as object `Common`. How many occupations have
salary data available from BLS?


```{r cache = TRUE}

 OccupationData %>% 
  left_join(Salaries, 
            by = c("O*NET-SOC Code","Title")) %>%
  select(Title, Median) %>%
  filter(!is.na(Median))

Common <- OccupationData %>% 
  left_join(Salaries, 
            by = c("O*NET-SOC Code","Title")) %>%
  filter(!is.na(Median))

  

```

Use basic `dplyr` verbs to answer each question.

<br>

**(2) Question:** Join new object `Common` with object `Interests`. Note that 
each interest is in `Element Name` and each interest rating (on a 7-point scale) 
is listed in `Data Value`. In effect, we can now associate each "interest" (e.g. 
artistic, social, enterprising) with median annual salaries.

Use the `lm()` function with formula `Median ~ Element Name` to print the 
coefficients of a simple linear model. Which two element names (not "High 
Points") are most associated with higher salaries?


```{r cache = TRUE}

options(scipen = 999)

dat <- Common %>% 
  left_join(Interests, 
            by = "O*NET-SOC Code") 

lm(formula=Median ~ `Element Name`,data=dat)


```
**Conventional** and **Enterprising** are most associated with higher salaries.
<br>

**(3) Question:** Now join `Common` with `TechnologySkills`. Filter occupations 
to only include the following technologies in `Example`, which have been 
provided for you in object `tech`: 

* R
* C
* Python
* JavaScript
* Microsoft Excel
* Spreadsheet software
* SAS statistical software
* Hypertext markup language HTML

Group by technologies (`Example`) and determine the following summaries:

* Average value for `Median` (annual salary)
* Count for total occupations with said `tech` 

```{r cache = TRUE}

tech <- c("R",
          "C++",
          "Python",
          "JavaScript",
          "Microsoft Excel",
          "Spreadsheet software",
          "SAS statistical software",
          "Hypertext markup language HTML")

Common %>% 
  left_join(TechnologySkills, 
            by = "O*NET-SOC Code") %>%
  filter(Example%in% tech) %>%
  group_by(Example) %>%
  summarize(Title=n(),
            Avg_med = mean(Median))

```

<br>

**(4) Question:** Please rate your own work values by providing a score from 
`1` to `7` for the following value categories.

* Achievement
* Working Conditions
* Recognition
* Relationships
* Support
* Independence 

A "7" indicates highest importance. A "1" indicates lowest importance. 
A default of `3.5` has been provided as an example.

```{r echo = TRUE}

my_values <- tibble::tibble("Achievement"        = 6, 
                            "Working Conditions" = 7,
                            "Recognition"        = 5,
                            "Relationships"      = 5,
                            "Support"            = 6,
                            "Independence"       = 5)

```

<br>

**(5) Question:** Join objects `Common` and `WorkValues` and filter the 
following values from variable `Element Name`:

* "First Work Value High-Point"
* "Second Work Value High-Point"
* "Third Work Value High-Point"

Then, use `select()` on the following variables:

* "O*NET-SOC Code"
* "Element Name" 
* "Data Value" 
* "Median"

Pipe (`%>%`) your output into the following `tidyr` function:

```{r eval = FALSE, include = TRUE}

work_values <- c("First Work Value High-Point",
"Second Work Value High-Point",
"Third Work Value High-Point")

 Common %>% 
  left_join(WorkValues, 
            by = "O*NET-SOC Code") %>%
  filter(`Element Name`%in% work_values) %>%
  select(`O*NET-SOC Code`,`Element Name`,`Data Value`,Median) %>%
  tidyr::pivot_wider(names_from = "Element Name", 
                   values_from = "Data Value")


```

<br>

Lastly, use function `lm()` and `predict()` to estimate your salary based on 
your work values! 

Make sure to set `eval = TRUE` for the following to run:

```{r eval = TRUE, include = TRUE, message = FALSE, warning = FALSE}

values <- c("Achievement", 
            "Working Conditions",
                            "Recognition",
                            "Relationships",
                            "Support",
                            "Independence" )

commons<- Common %>% 
  left_join(WorkValues, 
            by = "O*NET-SOC Code") %>%
  filter(`Element Name`%in% values) %>%
  select(`O*NET-SOC Code`,`Element Name`,`Data Value`,Median) %>%
  tidyr::pivot_wider(names_from = "Element Name", 
                   values_from = "Data Value")

my_model <- lm(formula = Median ~ 
                 Achievement +
                 `Working Conditions` +
                 Recognition + 
                 Relationships + 
                 Support + 
                 Independence, 
               data = commons)

summary(my_model)

predict(object = my_model, newdata = my_values)

```