---
title: "Lab 05: - Tempe Accident Analysis"
author: "Muge Kosar"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

<br>

# Setup

The following analysis uses data comprised of traffic accidents in the city of 
Tempe, Arizona. They constitute 28,470 observations and 32 variables. You can 
find more details in the dataset documentation at the 
[**Tempe Open Data Portal**](https://data.tempe.gov/dataset/high-severity-traffic-crashes-1-08). 

```{r setup, include = FALSE}

# SET UP GLOBAL OPTIONS

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10)

```

```{r include = FALSE}

# LOAD REQUIRED PACKAGES

library(ggmap)        # Packages per instructions
library(dplyr)
library(pander)
library(viridis)
library(ggplot2)
library(ggthemes)

library(scales)       # Recommended package

                      # Function 'percent()' from "scales" allows
                      # easy formatting of proportions to percentages

                      # Funtion 'number()' from "scales" allows 
                      # easy formatting of numbers with argument 
                      # 'big.mark = TRUE'. 

                      # Note that 'percent()' and 'number()' format 
                      # values as character strings! So they won't 
                      # work like numbers any more (e.g. in 'sum()')!

```

```{r include = FALSE}

# READ IN DATA

url <- paste0("https://github.com/DS4PS/Data-Science-Class/blob",
              "/master/DATA/TempeTrafficAccidents.rds?raw=true")

dat <- readRDS(gzcon(url(url)))     # Method per instructions

```

<br>

Explore the first six observations in the following interactive table:

```{r echo = TRUE}

head(dat)

```

<br>

# Preprocessing

All data preprocessing are available in **Appendix Code A: Data Preprocessing**.

```{r include = FALSE}

date.vec <- strptime(dat$DateTime, 
                     format = "%m/%d/%y %H:%M")   # Create date fields

dat$hour   <- format(date.vec, format = "%H")
dat$month  <- format(date.vec, format = "%b")
dat$day    <- format(date.vec, format = "%a")
dat$day365 <- format(date.vec, format = "%j")
dat$week   <- format(date.vec, format = "%V")

dat$day <- factor(dat$day, 
                  levels = c("Mon",
                             "Tue",
                             "Wed",
                             "Thu",
                             "Fri",
                             "Sat",
                             "Sun"))              # Order weekdays

dat$hour12 <- format(date.vec, 
                     format="%l %p")              # Create 12-hour format

time.levels <- c("12 AM", " 1 AM", " 2 AM", " 3 AM", " 4 AM", " 5 AM", 
                 " 6 AM", " 7 AM", " 8 AM", " 9 AM", "10 AM", "11 AM", 
                 "12 PM", " 1 PM", " 2 PM", " 3 PM", " 4 PM", " 5 PM", 
                 " 6 PM", " 7 PM", " 8 PM", " 9 PM", "10 PM", "11 PM" )

dat$hour12 <- factor(dat$hour12, 
                     levels = time.levels)        # Order time intervals

age.labels <- paste0("Age ", 
                     c(16,18,25,35,45,55,65,75), "-", 
                     c(18,25,35,45,55,65,75,100) )

dat$age <- cut(dat$Age_Drv1, 
               breaks = c(16,18,25,
                          35,45,55,
                          65,75,100), 
               labels = age.labels)               # Discretize age ranges

```

<br>

# Part I: Data Summaries

The following provides brief exploratory summaries on key data points.

<br>

## 1. Accidents on Mondays

```{r}

x <- sum(dat$day == "Mon", 
         na.rm = TRUE)

comma(x)                    # Example solution

```

**`r comma(x)`** accidents occured on Mondays.

<br>

## 2. Proportion of Accidents on Mondays

```{r}

x<- mean(dat$day == "Mon",
         na.rm = TRUE)
percent(x)

```

**`r percent(x)`** of all accidents occur on Mondays.

<br>

## 3. Proportion of Harmful Monday Accidents 

**Note:** *"Harm" is defiend as any accident that causes at least one injury or* 
*fatality, i.e. a casualty.*

```{r}

x <- dat %>%
  group_by(day) %>%
  filter(day == "Mon") %>%
  mutate(harm_acc = Totalinjuries > 0 | Totalfatalities > 0) %>%
  summarize(n = n(),
            harm.rate = mean(harm_acc))

percent(x$harm.rate)
  

```

**`r percent(x$harm.rate)`** of all Monday accidents have at least one casualty.

<br>

## 4. Most Common Monday Accident Type

```{r}

dat %>%
  filter(day == "Mon") %>%
  count(Collisionmanner) %>%
  arrange(desc(n))

```

**Rear End** collisions are the most common accident types.

<br>

## 5. Differences in Accidents

```{r}

dat %>%
  group_by(day) %>%
  mutate(harm_acc = Totalinjuries > 0 | Totalfatalities > 0) %>%
  summarize(n = n(),
            harm.rate = mean(harm_acc)) %>%
  arrange(desc(harm.rate))

```

Significantly more accidents occur **Thursday**.

<br>

## 6. Accidents by Weekday, Casualties, & Harm

```{r}

dat %>%
  group_by(day) %>%
  mutate(harm_acc = Totalinjuries > 0 | Totalfatalities > 0) %>%
  summarize(n = n(),
            injuries = sum(Totalinjuries),
            fatalities = sum(Totalfatalities),
            harm.rate = mean(harm_acc)) %>%
   arrange(desc(harm.rate))

```

<br>

# Part II: Age Groups

The following provides summaries of accidents by age groups.

<br>

## 1. Accidents by Hour & Age

```{r}

dat %>%
count(age, hour12) %>%
group_by(age) %>%
mutate(n.age = sum(n)) %>%
ungroup() %>%
group_by(hour12) %>%
mutate(n.hour = sum(n)) %>%
filter(hour12 == " 7 AM") %>%
arrange(desc(n.age))
  

```

Drivers of **Age 18-25** have the greatest number of accidents from 7:00 to 7:59 AM.

<br>

## 2. Proportion of Accidents by Hour & Age

```{r}

dat %>% 
  group_by(hour12, age) %>% 
  summarize(n = n()) %>%      
  group_by(age) %>%           
  mutate(n.age = mean(n)) %>%
  filter(hour12 == " 7 AM") %>%
  arrange(desc(n.age))


```

<br>

# Part III: Rates of Harm

The following reports the accidents, casualties, proportion, and average 
casualties per harmful accident.

<br>

## 1. Accidents by Hour

```{r}

dat %>%
  group_by(hour) %>%
  summarize(accidents = number(n(), 
                               big.mark = ","))     # Example solution

dat %>%
  group_by(hour) %>%
  summarize(n = n()) %>%
  plot(type = "b",
       bty = "n",
       pch = 19,
       cex = 2,
       xlab = "Hour",
       ylab = "Total Accidents",
       main = "Total Accidents by Time of Day")

```

<br>

## 2. Total Casualties by Hour

```{r}

dat %>%
  filter(Totalfatalities>0 | Totalinjuries >0) %>%
  group_by(hour) %>%
  summarize(n = n(),
            injuries = sum(Totalinjuries),
            fatalities = sum(Totalfatalities))

dat %>%
  filter(Totalfatalities>0 | Totalinjuries >0) %>%
  group_by(hour) %>%
  summarize(n = n()) %>%
  select(hour,n) %>%
  plot(type = "b",
       bty = "n",
       pch = 19,
       cex = 2,
       xlab = "Hour of the day",
       ylab = "Total Number of Passengers Hurt or Killed",
       main = "Total Injuries or Fatalities by Hour of the Day")
       

```

<br>

## 3. Proportion of Accidents with Casualties

```{r}
dat %>%
    group_by(hour) %>%
    summarize(Total_Harm = sum(Totalfatalities > 0 | Totalinjuries > 0),
              Ave = Total_Harm / n()) %>%
  select(hour,Ave) %>%
  plot(type = "b",
       bty = "n",
       pch = 19,
       cex = 2,
       xlab = "Hour of the day",
       ylab = "Proportion of Accidents Resulting in Harm",
       main = "Proportion of Crashes that Result in Injuries or Fatalities")

```

<br>

## 4. Average Casualties per Harmful Accident

```{r}

dat %>%
  filter(Totalfatalities>0 | Totalinjuries >0) %>%
  group_by(hour) %>%
    summarize(Total_Harm = sum(Totalfatalities, na.rm = TRUE) + sum(Totalinjuries, na.rm = TRUE),
              Ave = Total_Harm / n()) %>%
  select(hour, Ave) %>%
  plot(type = "b",
       bty = "n",
       pch = 19,
       cex = 2,
       xlab = "Hour of the day",
       ylab = "Ave. Number of Passengers Hurts",
       main = "Average Injuries or Fatalities per Harmful Crash")

```

<br>

# Appendix

<br>

## Code A: Data Preprocessing

```{r eval = FALSE, echo = TRUE, include = TRUE}

date.vec <- strptime(dat$DateTime, 
                     format = "%m/%d/%y %H:%M")   # Create date fields

dat$hour   <- format(date.vec, format = "%H")
dat$month  <- format(date.vec, format = "%b")
dat$day    <- format(date.vec, format = "%a")
dat$day365 <- format(date.vec, format = "%j")
dat$week   <- format(date.vec, format = "%V")

dat$day <- factor(dat$day, 
                  levels = c("Mon",
                             "Tue",
                             "Wed",
                             "Thu",
                             "Fri",
                             "Sat",
                             "Sun"))              # Order weekdays

dat$hour12 <- format(date.vec, 
                     format="%l %p")              # Create 12-hour format

time.levels <- c("12 AM", " 1 AM", " 2 AM", " 3 AM", " 4 AM", " 5 AM", 
                 " 6 AM", " 7 AM", " 8 AM", " 9 AM", "10 AM", "11 AM", 
                 "12 PM", " 1 PM", " 2 PM", " 3 PM", " 4 PM", " 5 PM", 
                 " 6 PM", " 7 PM", " 8 PM", " 9 PM", "10 PM", "11 PM" )

dat$hour12 <- factor(dat$hour12, 
                     levels = time.levels)        # Order time intervals

age.labels <- paste0("Age ", 
                     c(16,18,25,35,45,55,65,75), "-", 
                     c(18,25,35,45,55,65,75,100) )

dat$age <- cut(dat$Age_Drv1, 
               breaks = c(16,18,25,
                          35,45,55,
                          65,75,100), 
               labels = age.labels)               # Discretize age ranges

```

